---
title: "DATA 603 Project - Group 6"
date: "2025-03-28"
output: 
  pdf_document:
    latex_engine: xelatex
author: 
  - "Group 6"
---
1. INTRODUCTION

1.1 MOTIVATION

The housing market has been fairly inconsistent lately in various regions around the world. Prices have skyrocketed and it has become very difficult for young home owners to purchase homes. We are looking into this housing data in boston to gain an idea of what factors are considered in the valuation of houses and what prospective homeowners should look out for when trying to buy a house.
1.1.1 Context

This project would explore the relationships between housing prices in Boston and various features that may influence them.

1.1.2 Problem



1.1.3 Challenges

1.2 OBJECTIVES

1.2.1 Overview


1.2.2 Goals & Research Questions
Our goal is to develop a multiple linear regression model that accurately predicts house prices based on significant features.



2. METHODOLOGY
2.1 Data
2.2 Approach
2.3 Workflow

a. Multicolinearity
```{r}
ggplot(model, aes(x=.fitted, y=.resid)) +
geom_point() +geom_smooth()+
geom_hline(yintercept = 0)
# Example code for determining linearity
#if there is a pattern, then there is a linearity and we have to transform our data.

```

b. Additive model
```{r}
boston_data= read.csv("./BostonHousing.csv")
head(boston_data)

```
```{r}
boston_additive = lm(medv~crim+zn+indus+chas+nox+rm+age+dis+rad+tax+ptratio+b+lstat, data= boston_data)
summary(boston_additive)
```
c. Final Additive model (F-test, anova table)
Individual T test
stepwise
ols_backwards
ols_forward
Best Subset
HigherOrder
d. R squared & RSE
e. 
f. Interaction model
g.  

2.4 Contributions

3. MAIN RESULTS OF THE ANALYSIS
3.1 Results
Individual T-test: 

Hypothesis: 


Based off the values indus and age do not seem relevent. 

We will verify utilizing a globabal anova f-test: 
```{r}
boston_additive_dropped = lm(medv~crim+zn+chas+nox+rm+dis+rad+tax+ptratio+b+lstat, data= boston_data)
anova(boston_additive, boston_additive_dropped)

```
Fail to reject null, they are not relevent. 


Comparison Of R^2 and RSE 

Chas is a factor. 


Final Additive Model: 

```{r}
confint(boston_additive_dropped)
```

4. Conclusion and Discussion
4.1 Approach
4.2 Future Work.
