---
title: "DATA 603 Project - Group 6"
date: "2025-03-28"
output: 
  pdf_document:
    latex_engine: xelatex
author: 
  - "Group 6"
---
1. INTRODUCTION

1.1 MOTIVATION
1.1.1 Context

1.1.2 Problem

1.1.3 Challenges

1.2 OBJECTIVES
1.2.1 Overview
1.2.2 Goals & Research Questions


2. METHODOLOGY
2.1 Data
2.2 Approach
2.3 Workflow

a. Multicolinearity
b. Additive model
c. Final Additive model (F-test, anova table)
Individual T test
stepwise
ols_backwards
ols_forward
Best Subset
HigherOrder
d. R squared & RSE
e. 
f. Interaction model
g.  

2.4 Contributions

3. MAIN RESULTS OF THE ANALYSIS
3.1 Results
```{r}
boston_data= read.csv("./BostonHousing.csv")
head(boston_data)

```
```{r}
boston_additive = lm(medv~crim+zn+indus+chas+nox+rm+age+dis+rad+tax+ptratio+b+lstat, data= boston_data)
summary(boston_additive)
```
Individual T-test: 

Hypothesis: 


Based off the values indus and age do not seem relevent. 

We will verify utilizing a globabal anova f-test: 
```{r}
boston_additive_dropped = lm(medv~crim+zn+chas+nox+rm+dis+rad+tax+ptratio+b+lstat, data= boston_data)
anova(boston_additive, boston_additive_dropped)

```
Fail to reject null, they are not relevent. 


Comparison Of R^2 and RSE 

Chas is a factor. 


Final Additive Model: 

```{r}
confint(boston_additive_dropped)
```


